---
title: "Ordinal GAMuT Sims w/ Parents + Addtl Variant Stats"
author: "Taylor Fischer"
date: "6/5/2018"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r grabdirs, include=FALSE}
library(data.table)
library(reshape2)
library(kableExtra)
library(knitr)
library(dplyr)
library(ggplot2)
options(knitr.table.format = "html")

source("/Volumes/JFK_Passport/ordinal_family_history/code/results/helper_functions.r")

home_dir <- "/Volumes/JFK_Passport/ordinal_family_history/code/" ## include trailing slash
list_of_folders <- list.files(home_dir, pattern = "post")
```

## 1. Target Prevalence 0.01 vs. 0.05 (Lambda = 2)

```{r block1, echo=FALSE, warning=F, message=F, fig.height=8, fig.width=7}
input_dir <- c("ordinal_201803061259_post",
               "ordinal_201803081323_post",
               "ordinal_flagged_201803151437_post",
               "ordinal_flagged_201803290940_post",
               "ordinal_parent_sims_201805231339_post")
input_dir <- paste0(home_dir,input_dir,"/processed_data")
alpha <- 5e-5
filter_by <- list(Nsibsize=c(0,1,2), n_per_group=c(400,750,1000,1500), lambda=2)

list_of_files <- unlist(lapply(input_dir, list.files, full.names=TRUE))
l <- lapply(list_of_files, get_power, alpha=alpha)
dat <- {}
for(i in 1:length(l)) {dat <- rbind(dat,l[[i]])}
  
if(!is.na(filter_by)) {
    for(i in 1:length(filter_by)){
      varid <- names(filter_by[i])
      vals <- filter_by[[i]]
      dat <- dat[dat[,varid]%in%vals,]
    }
}
  
constant_params <- dat[1,apply(dat, 2, function(col) { length(unique(col))==1 })]
row.names(constant_params)<-"value"
dat$method <- factor(dat$method,levels=c("burden","skat","gamut"),
                     labels=c("Burden Test","SKAT","Ordinal GAMuT"))
dat$Nsibsize <- factor(dat$Nsibsize, levels=c(1,2,0), 
                       labels=c("N Siblings 1","N Siblings 2", "Parents Only"))
dat$effectsize <- factor(dat$effectsize, levels=c(2,4,6), 
                         labels=c("RV Effect Size 2","RV Effect Size 4","RV Effect Size 6"))
dat$lambda <- factor(dat$lambda)
dat$prev_target <- factor(dat$prev_target)

kable(t(constant_params), col.names=" ", caption="constant parameters") %>%
  kable_styling(bootstrap_options = c("condensed"), full_width = F, position="float_left") %>%
  row_spec(c(10), bold = T, color = "blue", background = "white")

ggplot(data = dat, aes_string(x = "n_per_group", y = "power", color="method", linetype="prev_target")) + 
  geom_line() + geom_point(alpha=0.6) + 
  scale_x_continuous(labels = unique(dat[,"n_per_group"]), breaks = unique(dat[,"n_per_group"])) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  scale_y_continuous(limits = c(0, 1), expand = c(0,0)) + 
  facet_grid(paste0("effectsize","~","Nsibsize")) + 
  guides(color=guide_legend(title="Method"),linetype=guide_legend(title="Prevalence")) +
  xlab("Samples/Group") + 
  ylab(paste0("Empirical Power (α = ",alpha,")"))

```

<P style="page-break-before: always">
## 2. Target Prevalence 0.01 vs. 0.05 (Lambda = 4)

```{r block2, echo=FALSE, warning=F, message=F, fig.height=8, fig.width=7}
input_dir <- c("ordinal_201803061259_post",
               "ordinal_201803081323_post",
               "ordinal_flagged_201803151437_post",
               "ordinal_flagged_201803290940_post",
               "ordinal_parent_sims_201805231339_post")
input_dir <- paste0(home_dir,input_dir,"/processed_data")
alpha <- 5e-5
filter_by <- list(Nsibsize=c(0,1,2), n_per_group=c(400,750,1000,1500), lambda=4)

list_of_files <- unlist(lapply(input_dir, list.files, full.names=TRUE))
l <- lapply(list_of_files, get_power, alpha=alpha)
dat <- {}
for(i in 1:length(l)) {dat <- rbind(dat,l[[i]])}
  
if(!is.na(filter_by)) {
    for(i in 1:length(filter_by)){
      varid <- names(filter_by[i])
      vals <- filter_by[[i]]
      dat <- dat[dat[,varid]%in%vals,]
    }
}
  
constant_params <- dat[1,apply(dat, 2, function(col) { length(unique(col))==1 })]
row.names(constant_params)<-"value"
dat$method <- factor(dat$method,levels=c("burden","skat","gamut"),
                     labels=c("Burden Test","SKAT","Ordinal GAMuT"))
dat$Nsibsize <- factor(dat$Nsibsize, levels=c(1,2,0), 
                       labels=c("N Siblings 1","N Siblings 2", "Parents Only"))
dat$effectsize <- factor(dat$effectsize, levels=c(2,4,6), 
                         labels=c("RV Effect Size 2","RV Effect Size 4","RV Effect Size 6"))
dat$lambda <- factor(dat$lambda)
dat$prev_target <- factor(dat$prev_target)

kable(t(constant_params), col.names=" ", caption="constant parameters") %>%
  kable_styling(bootstrap_options = c("condensed"), full_width = F, position="float_left") %>%
  row_spec(c(10), bold = T, color = "blue", background = "white")

ggplot(data = dat, aes_string(x = "n_per_group", y = "power", color="method", linetype="prev_target")) + 
  geom_line() + geom_point(alpha=0.6) + 
  scale_x_continuous(labels = unique(dat[,"n_per_group"]), breaks = unique(dat[,"n_per_group"])) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  scale_y_continuous(limits = c(0, 1), expand = c(0,0)) + 
  facet_grid(paste0("effectsize","~","Nsibsize")) + 
  guides(color=guide_legend(title="Method"),linetype=guide_legend(title="Prevalence")) +
  xlab("Samples/Group") + 
  ylab(paste0("Empirical Power (α = ",alpha,")"))
```

<P style="page-break-before: always">
## 3. Target Prevalence 0.01 vs. 0.05 (Lambda = 8)

```{r block3, echo=FALSE, warning=F, message=F, fig.height=8, fig.width=7}
input_dir <- c("ordinal_201803061259_post",
               "ordinal_201803081323_post",
               "ordinal_flagged_201803151437_post",
               "ordinal_flagged_201803290940_post",
               "ordinal_parent_sims_201805231339_post")
input_dir <- paste0(home_dir,input_dir,"/processed_data")
alpha <- 5e-5
filter_by <- list(Nsibsize=c(0,1,2), n_per_group=c(400,750,1000,1500), lambda=8)

list_of_files <- unlist(lapply(input_dir, list.files, full.names=TRUE))
l <- lapply(list_of_files, get_power, alpha=alpha)
dat <- {}
for(i in 1:length(l)) {dat <- rbind(dat,l[[i]])}
  
if(!is.na(filter_by)) {
    for(i in 1:length(filter_by)){
      varid <- names(filter_by[i])
      vals <- filter_by[[i]]
      dat <- dat[dat[,varid]%in%vals,]
    }
}
  
constant_params <- dat[1,apply(dat, 2, function(col) { length(unique(col))==1 })]
row.names(constant_params)<-"value"
dat$method <- factor(dat$method,levels=c("burden","skat","gamut"),
                     labels=c("Burden Test","SKAT","Ordinal GAMuT"))
dat$Nsibsize <- factor(dat$Nsibsize, levels=c(1,2,0), 
                       labels=c("N Siblings 1","N Siblings 2","Parents Only"))
dat$effectsize <- factor(dat$effectsize, levels=c(2,4,6), 
                         labels=c("RV Effect Size 2","RV Effect Size 4","RV Effect Size 6"))
dat$lambda <- factor(dat$lambda)
dat$prev_target <- factor(dat$prev_target)

kable(t(constant_params), col.names=" ", caption="constant parameters") %>%
  kable_styling(bootstrap_options = c("condensed"), full_width = F, position="float_left") %>%
  row_spec(c(10), bold = T, color = "blue", background = "white")

ggplot(data = dat, aes_string(x = "n_per_group", y = "power", color="method", linetype="prev_target")) + 
  geom_line() + geom_point(alpha=0.6) + 
  scale_x_continuous(labels = unique(dat[,"n_per_group"]), breaks = unique(dat[,"n_per_group"])) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  scale_y_continuous(limits = c(0, 1), expand = c(0,0)) + 
  facet_grid(paste0("effectsize","~","Nsibsize")) + 
  guides(color=guide_legend(title="Method"),linetype=guide_legend(title="Prevalence")) +
  xlab("Samples/Group") + 
  ylab(paste0("Empirical Power (α = ",alpha,")"))

```

<P style="page-break-before: always">

## 4. Target Prevalence 0.01 vs. 0.05 (RV Effect Size = 2)

***NOTE: For figures 4-6 below, the same data as above is represented with different axes.

```{r block4, echo=FALSE, warning=F, message=F, fig.height=8, fig.width=7}
input_dir <- c("ordinal_201803061259_post",
               "ordinal_201803081323_post",
               "ordinal_flagged_201803151437_post",
               "ordinal_flagged_201803290940_post",
               "ordinal_parent_sims_201805231339_post")
input_dir <- paste0(home_dir,input_dir,"/processed_data")
alpha <- 5e-2
filter_by <- list(Nsibsize=c(0,1,2), n_per_group=c(400,750,1000,1500), effectsize=2)

list_of_files <- unlist(lapply(input_dir, list.files, full.names=TRUE))
l <- lapply(list_of_files, get_power, alpha=alpha)
dat <- {}
for(i in 1:length(l)) {dat <- rbind(dat,l[[i]])}
  
if(!is.na(filter_by)) {
    for(i in 1:length(filter_by)){
      varid <- names(filter_by[i])
      vals <- filter_by[[i]]
      dat <- dat[dat[,varid]%in%vals,]
    }
}
  
constant_params <- dat[1,apply(dat, 2, function(col) { length(unique(col))==1 })]
row.names(constant_params)<-"value"
dat$method <- factor(dat$method,levels=c("burden","skat","gamut"),
                     labels=c("Burden Test","SKAT","Ordinal GAMuT"))
dat$Nsibsize <- factor(dat$Nsibsize, levels=c(1,2,0), 
                       labels=c("N Siblings 1","N Siblings 2","Parents Only"))
dat$lambda <- factor(dat$lambda, levels=c(2,4,8), 
                         labels=c("Lambda 2","Lambda 4","Lambda 8"))
dat$prev_target <- factor(dat$prev_target)

kable(t(constant_params), col.names=" ", caption="constant parameters") %>%
  kable_styling(bootstrap_options = c("condensed"), full_width = F, position="float_left") %>%
  row_spec(c(10), bold = T, color = "blue", background = "white")

ggplot(data = dat, aes_string(x = "n_per_group", y = "power", color="method", linetype="prev_target")) + 
  geom_line() + geom_point(alpha=0.6) + 
  scale_x_continuous(labels = unique(dat[,"n_per_group"]), breaks = unique(dat[,"n_per_group"])) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  scale_y_continuous(limits = c(0, 1), expand = c(0,0)) + 
  facet_grid(paste0("Nsibsize","~","lambda")) + 
  guides(color=guide_legend(title="Method"),linetype=guide_legend(title="Prevalence")) +
  xlab("Samples/Group") + 
  ylab(paste0("Empirical Power (α = ",alpha,")"))

```

<P style="page-break-before: always">
## 5. Target Prevalence 0.01 vs. 0.05 (RV Effect Size = 4)

```{r block5, echo=FALSE, warning=F, message=F, fig.height=8, fig.width=7}
input_dir <- c("ordinal_201803061259_post",
               "ordinal_201803081323_post",
               "ordinal_flagged_201803151437_post",
               "ordinal_flagged_201803290940_post",
               "ordinal_parent_sims_201805231339_post")
input_dir <- paste0(home_dir,input_dir,"/processed_data")
alpha <- 5e-5
filter_by <- list(Nsibsize=c(0,1,2), n_per_group=c(400,750,1000,1500), effectsize=4)

list_of_files <- unlist(lapply(input_dir, list.files, full.names=TRUE))
l <- lapply(list_of_files, get_power, alpha=alpha)
dat <- {}
for(i in 1:length(l)) {dat <- rbind(dat,l[[i]])}
  
if(!is.na(filter_by)) {
    for(i in 1:length(filter_by)){
      varid <- names(filter_by[i])
      vals <- filter_by[[i]]
      dat <- dat[dat[,varid]%in%vals,]
    }
}
  
constant_params <- dat[1,apply(dat, 2, function(col) { length(unique(col))==1 })]
row.names(constant_params)<-"value"
dat$method <- factor(dat$method,levels=c("burden","skat","gamut"),
                     labels=c("Burden Test","SKAT","Ordinal GAMuT"))
dat$Nsibsize <- factor(dat$Nsibsize, levels=c(1,2,0), 
                       labels=c("N Siblings 1","N Siblings 2","Parents Only"))
dat$lambda <- factor(dat$lambda, levels=c(2,4,8), 
                         labels=c("Lambda 2","Lambda 4","Lambda 8"))
dat$prev_target <- factor(dat$prev_target)

kable(t(constant_params), col.names=" ", caption="constant parameters") %>%
  kable_styling(bootstrap_options = c("condensed"), full_width = F, position="float_left") %>%
  row_spec(c(10), bold = T, color = "blue", background = "white")

ggplot(data = dat, aes_string(x = "n_per_group", y = "power", color="method", linetype="prev_target")) + 
  geom_line() + geom_point(alpha=0.6) + 
  scale_x_continuous(labels = unique(dat[,"n_per_group"]), breaks = unique(dat[,"n_per_group"])) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  scale_y_continuous(limits = c(0, 1), expand = c(0,0)) + 
  facet_grid(paste0("Nsibsize","~","lambda")) + 
  guides(color=guide_legend(title="Method"),linetype=guide_legend(title="Prevalence")) +
  xlab("Samples/Group") + 
  ylab(paste0("Empirical Power (α = ",alpha,")"))

```

<P style="page-break-before: always">
## 6. Target Prevalence 0.01 vs. 0.05 (RV Effect Size = 6)

```{r block6, echo=FALSE, warning=F, message=F, fig.height=8, fig.width=7}
input_dir <- c("ordinal_201803061259_post",
               "ordinal_201803081323_post",
               "ordinal_flagged_201803151437_post",
               "ordinal_flagged_201803290940_post",
               "ordinal_parent_sims_201805231339_post")
input_dir <- paste0(home_dir,input_dir,"/processed_data")
alpha <- 5e-5
filter_by <- list(Nsibsize=c(0,1,2), n_per_group=c(400,750,1000,1500), effectsize=6)

list_of_files <- unlist(lapply(input_dir, list.files, full.names=TRUE))
l <- lapply(list_of_files, get_power, alpha=alpha)
dat <- {}
for(i in 1:length(l)) {dat <- rbind(dat,l[[i]])}
  
if(!is.na(filter_by)) {
    for(i in 1:length(filter_by)){
      varid <- names(filter_by[i])
      vals <- filter_by[[i]]
      dat <- dat[dat[,varid]%in%vals,]
    }
}
  
constant_params <- dat[1,apply(dat, 2, function(col) { length(unique(col))==1 })]
row.names(constant_params)<-"value"
dat$method <- factor(dat$method,levels=c("burden","skat","gamut"),
                     labels=c("Burden Test","SKAT","Ordinal GAMuT"))
dat$Nsibsize <- factor(dat$Nsibsize, levels=c(1,2,0), 
                       labels=c("N Siblings 1","N Siblings 2","Parents Only"))
dat$lambda <- factor(dat$lambda, levels=c(2,4,8), 
                         labels=c("Lambda 2","Lambda 4","Lambda 8"))
dat$prev_target <- factor(dat$prev_target)

kable(t(constant_params), col.names=" ", caption="constant parameters") %>%
  kable_styling(bootstrap_options = c("condensed"), full_width = F, position="float_left") %>%
  row_spec(c(10), bold = T, color = "blue", background = "white")

ggplot(data = dat, aes_string(x = "n_per_group", y = "power", color="method", linetype="prev_target")) + 
  geom_line() + geom_point(alpha=0.6) + 
  scale_x_continuous(labels = unique(dat[,"n_per_group"]), breaks = unique(dat[,"n_per_group"])) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  scale_y_continuous(limits = c(0, 1), expand = c(0,0)) + 
  facet_grid(paste0("Nsibsize","~","lambda")) + 
  guides(color=guide_legend(title="Method"),linetype=guide_legend(title="Prevalence")) +
  xlab("Samples/Group") + 
  ylab(paste0("Empirical Power (α = ",alpha,")"))

```


<P style="page-break-before: always">
## 7. Variant Stats (Effect Size=2, N Per Group=1000)

***NOTE: The results below apply only to simulations with 1 or 2 siblings. No simulation data with parents included as family history are shown. 

```{r block8_2, echo=FALSE, warning=F, message=F, fig.height=5, fig.width=10}
input_dir <- c("ordinal_flagged_201803141731_post",
                "ordinal_flagged_201803151437_post",
               "ordinal_flagged_201803290940_post",
               "ordinal_flagged_201805241446_post")
load(paste0(home_dir,input_dir[1],"/processed_flagged_data/flagged.RData"))
idat <- final
load(paste0(home_dir,input_dir[2],"/processed_flagged_data/flagged.RData"))
idat <- rbind(idat, final)
load(paste0(home_dir,input_dir[3],"/processed_flagged_data/flagged.RData"))
idat <- rbind(idat, final)
load(paste0(home_dir,input_dir[4],"/processed_flagged_data/flagged.RData"))
idat <- rbind(idat, final)


idat <- idat[idat$Nsibsize%in%c(1,2),]
idat <- idat[idat$n_per_group%in%1000,]
idat <- idat[idat$effectsize%in%2,]

idat$group <- as.character(idat$group)
idat$group <- factor(idat$group, levels=c("0_kept=","1_kept=","2_kept="),
                      labels=c("Control",
                               "Case NFH",
                               "Case FH"))

idat$Nsibsize <- factor(idat$Nsibsize,levels=c(1,2),labels=c("N Siblings 1", "N Siblings 2"))
idat$lambda <- factor(idat$lambda, levels=c(2,4,8), labels=c("Lambda 2", "Lambda 4", "Lambda 8"))
idat$prev_target <- factor(idat$prev_target)

# avg MAF
plotdat <- idat[idat$measure%in%"avg.maf.IP",]
ggplot(data=plotdat,aes(x=group,y=mean,fill=prev_target))+
  geom_bar(color="black",stat="identity",position = "dodge")+ggtitle("Average MAF in index proband")+
  facet_grid(Nsibsize~lambda)+labs(y="Average (of 1000 Simulations)",x="Group")+
 # theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(), axis.title.x=element_blank())+
  guides(fill = guide_legend("Target Prev"))+
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                position=position_dodge(.9))+ 
  geom_text(
    aes(label = round(mean,3), y = mean+sd+ 0.002),
    position = position_dodge(0.9),
    vjust = 0, size=3
  )

# n.variants.IP
plotdat <- idat[idat$measure%in%"n.variants.IP",]
ggplot(data=plotdat,aes(x=group,y=mean,fill=prev_target))+
  geom_bar(color="black",stat="identity",position = "dodge")+ggtitle("Average number of variants in index proband")+
  facet_grid(Nsibsize~lambda)+labs(y="Average (of 1000 Simulations)",x="Group")+
 # theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(), axis.title.x=element_blank())+
  guides(fill = guide_legend("Target Prev"))+
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                position=position_dodge(.9))+ 
  geom_text(
    aes(label = round(mean,2), y = mean+sd+ 1),
    position = position_dodge(0.9),
    vjust = 0, size=3
  )

# n rare variants
plotdat <- idat[idat$measure%in%"n.rare.IP",]
ggplot(data=plotdat,aes(x=group,y=mean,fill=prev_target))+
  geom_bar(color="black",stat="identity",position = "dodge")+ggtitle("Average number of rare variants in index proband")+
  facet_grid(Nsibsize~lambda)+labs(y="Average (of 1000 Simulations)",x="Group")+
 # theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(), axis.title.x=element_blank())+
  guides(fill = guide_legend("Target Prev"))+
  geom_errorbar(aes(ymin=mean, ymax=mean+sd), width=.2,
                position=position_dodge(.9))+ 
  geom_text(
    aes(label = round(mean,2), y = mean+sd+0.08),
    position = position_dodge(0.9),
    vjust = 0, size=3
  )

# n causal variants
plotdat <- idat[idat$measure%in%"n.causal.IP",]
ggplot(data=plotdat,aes(x=group,y=mean,fill=prev_target))+
  geom_bar(color="black",stat="identity",position = "dodge")+ggtitle("Average number of causal rare variants in index proband")+
  facet_grid(Nsibsize~lambda)+labs(y="Average (of 1000 Simulations)",x="Group")+
 # theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(), axis.title.x=element_blank())+
  guides(fill = guide_legend("Target Prev"))+
  geom_errorbar(aes(ymin=mean, ymax=mean+sd), width=.2,
                position=position_dodge(.9))+ 
  geom_text(
    aes(label = round(mean,2), y = mean+sd+0.08),
    position = position_dodge(0.9),
    vjust = 0, size=3
  )

# prob disease
plotdat <- idat[idat$measure%in%"pdis.IP",]
ggplot(data=plotdat,aes(x=group,y=mean,fill=prev_target))+
  geom_bar(color="black",stat="identity",position = "dodge")+ggtitle("Average probability of disease of index proband")+
  facet_grid(Nsibsize~lambda)+labs(y="Average (of 1000 Simulations)",x="Group")+
 # theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(), axis.title.x=element_blank())+
  guides(fill = guide_legend("Target Prev"))+
  geom_errorbar(aes(ymin=mean, ymax=mean+sd), width=.2,
                position=position_dodge(.9))+ 
  geom_text(
    aes(label = round(mean,2), y = mean+sd+0.015),
    position = position_dodge(0.9),
    vjust = 0, size=3
  )
```