---
title: "Modified script from Taylor's for power plots for manuscript"
author: "CSL"
date: "1/22/2019"
output: html_document
---

We had to copy specific files:
```
cd Dropbox/Documents/gwas/projects/ordinal-skat/
scp csolislemus@hgcc.genetics.emory.edu:/home/csolislemus/ordinal_family_history_files_for_claudia/simulation_code/results/helper_functions.r scripts
```

We also need to have the following input directories:
```
input_dir <- c("ordinal_201803061259_post",
               "ordinal_201803081323_post",
               "ordinal_flagged_201803151437_post",
               "ordinal_flagged_201803290940_post",
               "ordinal_parent_sims_201805231339_post")
```

But in `simulation_code` folder, we only have:
```
hgcc:node00:[simulation_code] % ls
index.xlsx
instructions.txt
ordinal_201803061259_post.zip
ordinal_201803081323_post.zip
ordinal_error_201806061014_tmp.zip
ordinal_error_files_for_andrei_20180710_post
ordinal_error_files_for_andrei_20180710_post.zip
ordinal_flagged_201803121339_post.zip
ordinal_flagged_201805241446_post.zip
ordinal_parent_sims_201805231339_post.zip
results
```

So, we are missing the "flagged" ones. I emailed Taylor, and she sent them to me via Box.
I will copy them to hgcc (as backup):
```
cd Downloads
scp ordinal_flagged* csolislemus@hgcc.genetics.emory.edu:/home/csolislemus/ordinal_family_history_files_for_claudia/simulation_code
```

**Note** that we will run everything in hgcc, so not locally. This means that we will not run this script in its entirety, just copying line by line in the terminal.

We need to unzip the input folders:
```
input_dir <- c("ordinal_201803061259_post",
               "ordinal_201803081323_post",
               "ordinal_flagged_201803151437_post",
               "ordinal_flagged_201803290940_post",
               "ordinal_parent_sims_201805231339_post")
```

So
```
unzip ordinal_201803061259_post.zip
unzip ordinal_201803081323_post.zip
unzip ordinal_flagged_201803151437_post.zip
unzip ordinal_flagged_201803290940_post.zip
unzip ordinal_parent_sims_201805231339_post.zip
```

Now, we login hgcc:
```
ssh csolislemus@hgcc.genetics.emory.edu
qlogin -q i.q
module load R
```

<!--
We are not running as knit
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
-->


```{r grabdirs, include=FALSE}
library(data.table)
library(reshape2)
library(kableExtra)
##library(knitr)
library(dplyr)
library(ggplot2)
##options(knitr.table.format = "html")

source("/home/csolislemus/ordinal_family_history_files_for_claudia/simulation_code/results/helper_functions.r")

home_dir <- "/home/csolislemus/ordinal_family_history_files_for_claudia/simulation_code/" ## include trailing slash
##list_of_folders <- list.files(home_dir, pattern = "post")
```

We do not care about the plots 1-3 "Target Prevalence 0.01 vs. 0.05 (Lambda = 2,4,8)" as we want to make power plots with lambda as columns and prevalence as rows, one per RV effect size.

## 4. Target Prevalence 0.01 vs. 0.05 (RV Effect Size = 2)


```{r block4, echo=FALSE, warning=F, message=F, fig.height=8, fig.width=7}
input_dir <- c("ordinal_201803061259_post",
               "ordinal_201803081323_post",
               "ordinal_flagged_201803151437_post",
               "ordinal_flagged_201803290940_post",
               "ordinal_parent_sims_201805231339_post")
input_dir <- paste0(home_dir,input_dir,"/processed_data")
alpha <- 5e-2
filter_by <- list(Nsibsize=c(0,1,2), n_per_group=c(400,750,1000,1500), effectsize=2)

list_of_files <- unlist(lapply(input_dir, list.files, full.names=TRUE))
l <- lapply(list_of_files, get_power, alpha=alpha)
dat <- {}
for(i in 1:length(l)) {dat <- rbind(dat,l[[i]])}
  
if(!is.na(filter_by)) {
    for(i in 1:length(filter_by)){
      varid <- names(filter_by[i])
      vals <- filter_by[[i]]
      dat <- dat[dat[,varid]%in%vals,]
    }
}
  
constant_params <- dat[1,apply(dat, 2, function(col) { length(unique(col))==1 })]
row.names(constant_params)<-"value"
dat$method <- factor(dat$method,levels=c("burden","skat","gamut"),
                     labels=c("Burden Test","SKAT","Ordinal GAMuT"))
dat$Nsibsize <- factor(dat$Nsibsize, levels=c(1,2,0), 
                       labels=c("N Siblings 1","N Siblings 2","Parents Only"))
dat$lambda <- factor(dat$lambda, levels=c(2,4,8), 
                         labels=c("Lambda 2","Lambda 4","Lambda 8"))
dat$prev_target <- factor(dat$prev_target)

##kable(t(constant_params), col.names=" ", caption="constant parameters") %>%
##  kable_styling(bootstrap_options = c("condensed"), full_width = F, ##position="float_left") %>%
##  row_spec(c(10), bold = T, color = "blue", background = "white")

## We will write the dataframe to get the plots locally
write.table(dat,file="target-prevalence-rv2.csv", row.names = FALSE)
```

## 5. Target Prevalence 0.01 vs. 0.05 (RV Effect Size = 4)

```{r block5, echo=FALSE, warning=F, message=F, fig.height=8, fig.width=7}
input_dir <- c("ordinal_201803061259_post",
               "ordinal_201803081323_post",
               "ordinal_flagged_201803151437_post",
               "ordinal_flagged_201803290940_post",
               "ordinal_parent_sims_201805231339_post")
input_dir <- paste0(home_dir,input_dir,"/processed_data")
alpha <- 5e-5
filter_by <- list(Nsibsize=c(0,1,2), n_per_group=c(400,750,1000,1500), effectsize=4)

list_of_files <- unlist(lapply(input_dir, list.files, full.names=TRUE))
l <- lapply(list_of_files, get_power, alpha=alpha)
dat <- {}
for(i in 1:length(l)) {dat <- rbind(dat,l[[i]])}
  
if(!is.na(filter_by)) {
    for(i in 1:length(filter_by)){
      varid <- names(filter_by[i])
      vals <- filter_by[[i]]
      dat <- dat[dat[,varid]%in%vals,]
    }
}
  
constant_params <- dat[1,apply(dat, 2, function(col) { length(unique(col))==1 })]
row.names(constant_params)<-"value"
dat$method <- factor(dat$method,levels=c("burden","skat","gamut"),
                     labels=c("Burden Test","SKAT","Ordinal GAMuT"))
dat$Nsibsize <- factor(dat$Nsibsize, levels=c(1,2,0), 
                       labels=c("N Siblings 1","N Siblings 2","Parents Only"))
dat$lambda <- factor(dat$lambda, levels=c(2,4,8), 
                         labels=c("Lambda 2","Lambda 4","Lambda 8"))
dat$prev_target <- factor(dat$prev_target)

##kable(t(constant_params), col.names=" ", caption="constant parameters") %>%
##  kable_styling(bootstrap_options = c("condensed"), full_width = F, ##position="float_left") %>%
##  row_spec(c(10), bold = T, color = "blue", background = "white")

## We will write the dataframe to get the plots locally
write.table(dat,file="target-prevalence-rv4.csv", row.names = FALSE)
```


## 6. Target Prevalence 0.01 vs. 0.05 (RV Effect Size = 6)

```{r block6, echo=FALSE, warning=F, message=F, fig.height=8, fig.width=7}
input_dir <- c("ordinal_201803061259_post",
               "ordinal_201803081323_post",
               "ordinal_flagged_201803151437_post",
               "ordinal_flagged_201803290940_post",
               "ordinal_parent_sims_201805231339_post")
input_dir <- paste0(home_dir,input_dir,"/processed_data")
alpha <- 5e-5
filter_by <- list(Nsibsize=c(0,1,2), n_per_group=c(400,750,1000,1500), effectsize=6)

list_of_files <- unlist(lapply(input_dir, list.files, full.names=TRUE))
l <- lapply(list_of_files, get_power, alpha=alpha)
dat <- {}
for(i in 1:length(l)) {dat <- rbind(dat,l[[i]])}
  
if(!is.na(filter_by)) {
    for(i in 1:length(filter_by)){
      varid <- names(filter_by[i])
      vals <- filter_by[[i]]
      dat <- dat[dat[,varid]%in%vals,]
    }
}
  
constant_params <- dat[1,apply(dat, 2, function(col) { length(unique(col))==1 })]
row.names(constant_params)<-"value"
dat$method <- factor(dat$method,levels=c("burden","skat","gamut"),
                     labels=c("Burden Test","SKAT","Ordinal GAMuT"))
dat$Nsibsize <- factor(dat$Nsibsize, levels=c(1,2,0), 
                       labels=c("N Siblings 1","N Siblings 2","Parents Only"))
dat$lambda <- factor(dat$lambda, levels=c(2,4,8), 
                         labels=c("Lambda 2","Lambda 4","Lambda 8"))
dat$prev_target <- factor(dat$prev_target)

##kable(t(constant_params), col.names=" ", caption="constant parameters") %>%
##  kable_styling(bootstrap_options = c("condensed"), full_width = F, ##position="float_left") %>%
##  row_spec(c(10), bold = T, color = "blue", background = "white")

## We will write the dataframe to get the plots locally
write.table(dat,file="target-prevalence-rv6.csv", row.names = FALSE)
```


Now, we want to copy the data frames back:
```
cd Dropbox/Documents/gwas/projects/ordinal-skat/results
scp csolislemus@hgcc.genetics.emory.edu:/home/csolislemus/ordinal_family_history_files_for_claudia/simulation_code/*.csv .
```

We want to put the following plot (n causal variants) in the manuscript (or all? also probability of disease one).

## 7. Variant Stats (Effect Size=2, N Per Group=1000)

***NOTE: The results below apply only to simulations with 1 or 2 siblings. No simulation data with parents included as family history are shown. 

We do not care about the first 3 plots: average MAF, average number variants and average number of rare variants, so I removed them.

First, we need to unzip another folder in hgcc:
```
unzip ordinal_flagged_201805241446_post.zip
```

Taylor emailed the missing folder `ordinal_flagged_201803141731_post`, which we will copy to hgcc and unzip:
```
cd Downloads
scp ordinal_flagged_201803141731_post.zip csolislemus@hgcc.genetics.emory.edu:/home/csolislemus/ordinal_family_history_files_for_claudia/simulation_code/

## Then in HGCC:
unzip ordinal_flagged_201803141731_post.zip
```

Now, we login hgcc:
```
ssh csolislemus@hgcc.genetics.emory.edu
qlogin -q i.q
module load R
```

Now, we create the csv file that we will use later in plots:

```{r block8_2, echo=FALSE, warning=F, message=F, fig.height=5, fig.width=10}
input_dir <- c("ordinal_flagged_201803141731_post",
                "ordinal_flagged_201803151437_post",
               "ordinal_flagged_201803290940_post",
               "ordinal_flagged_201805241446_post")
load(paste0(home_dir,input_dir[1],"/processed_flagged_data/flagged.RData"))
idat <- final
load(paste0(home_dir,input_dir[2],"/processed_flagged_data/flagged.RData"))
idat <- rbind(idat, final)
load(paste0(home_dir,input_dir[3],"/processed_flagged_data/flagged.RData"))
idat <- rbind(idat, final)
load(paste0(home_dir,input_dir[4],"/processed_flagged_data/flagged.RData"))
idat <- rbind(idat, final)


idat <- idat[idat$Nsibsize%in%c(1,2),]
idat <- idat[idat$n_per_group%in%1000,]
idat <- idat[idat$effectsize%in%2,]

idat$group <- as.character(idat$group)
idat$group <- factor(idat$group, levels=c("0_kept=","1_kept=","2_kept="),
                      labels=c("Control",
                               "Case NFH",
                               "Case FH"))

idat$Nsibsize <- factor(idat$Nsibsize,levels=c(1,2),labels=c("N Siblings 1", "N Siblings 2"))
idat$lambda <- factor(idat$lambda, levels=c(2,4,8), labels=c("Lambda 2", "Lambda 4", "Lambda 8"))
idat$prev_target <- factor(idat$prev_target)


write.table(idat, file="variant-stats.csv", sep=",", row.names=FALSE, quote = FALSE)
```

Now, we want to copy the data frames back:
```
cd Dropbox/Documents/gwas/projects/ordinal-skat/results
scp csolislemus@hgcc.genetics.emory.edu:/home/csolislemus/ordinal_family_history_files_for_claudia/simulation_code/variant-stats.csv .
```


# Plots

These are the old plots. We will copy and create a modified version for the manuscript.
```
dat = read.table("target-prevalence-rv2.csv", header=TRUE)
alpha <- 5e-2
dat=within(dat,prev_target<-factor(prev_target))
ggplot(data = dat, aes_string(x = "n_per_group", y = "power", color="method", linetype="prev_target")) + 
  geom_line() + geom_point(alpha=0.6) + 
  scale_x_continuous(labels = unique(dat[,"n_per_group"]), breaks = unique(dat[,"n_per_group"])) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  scale_y_continuous(limits = c(0, 1), expand = c(0,0)) + 
  facet_grid(paste0("Nsibsize","~","lambda")) + 
  guides(color=guide_legend(title="Method"),linetype=guide_legend(title="Prevalence")) +
  xlab("Samples/Group") + 
  ylab(paste0("Empirical Power (α = ",alpha,")"))
```

```
dat = read.table("target-prevalence-rv4.csv", header=TRUE)
alpha <- 5e-5
ggplot(data = dat, aes_string(x = "n_per_group", y = "power", color="method", linetype="prev_target")) + 
  geom_line() + geom_point(alpha=0.6) + 
  scale_x_continuous(labels = unique(dat[,"n_per_group"]), breaks = unique(dat[,"n_per_group"])) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  scale_y_continuous(limits = c(0, 1), expand = c(0,0)) + 
  facet_grid(paste0("Nsibsize","~","lambda")) + 
  guides(color=guide_legend(title="Method"),linetype=guide_legend(title="Prevalence")) +
  xlab("Samples/Group") + 
  ylab(paste0("Empirical Power (α = ",alpha,")"))
```

```
dat = read.table("target-prevalence-rv6.csv", header=TRUE)
alpha <- 5e-5
ggplot(data = dat, aes_string(x = "n_per_group", y = "power", color="method", linetype="prev_target")) + 
  geom_line() + geom_point(alpha=0.6) + 
  scale_x_continuous(labels = unique(dat[,"n_per_group"]), breaks = unique(dat[,"n_per_group"])) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  scale_y_continuous(limits = c(0, 1), expand = c(0,0)) + 
  facet_grid(paste0("Nsibsize","~","lambda")) + 
  guides(color=guide_legend(title="Method"),linetype=guide_legend(title="Prevalence")) +
  xlab("Samples/Group") + 
  ylab(paste0("Empirical Power (α = ",alpha,")"))
```


```
idat = read.table("variant-stats.csv", header=TRUE)

# n causal variants

plotdat <- idat[idat$measure%in%"n.causal.IP",]
ggplot(data=plotdat,aes(x=group,y=mean,fill=prev_target))+
  geom_bar(color="black",stat="identity",position = "dodge")+ggtitle("Average number of causal rare variants in index proband")+
  facet_grid(Nsibsize~lambda)+labs(y="Average (of 1000 Simulations)",x="Group")+
 # theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(), axis.title.x=element_blank())+
  guides(fill = guide_legend("Target Prev"))+
  geom_errorbar(aes(ymin=mean, ymax=mean+sd), width=.2,
                position=position_dodge(.9))+ 
  geom_text(
    aes(label = round(mean,2), y = mean+sd+0.08),
    position = position_dodge(0.9),
    vjust = 0, size=3
  )

# prob disease
plotdat <- idat[idat$measure%in%"pdis.IP",]
ggplot(data=plotdat,aes(x=group,y=mean,fill=prev_target))+
  geom_bar(color="black",stat="identity",position = "dodge")+ggtitle("Average probability of disease of index proband")+
  facet_grid(Nsibsize~lambda)+labs(y="Average (of 1000 Simulations)",x="Group")+
 # theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(), axis.title.x=element_blank())+
  guides(fill = guide_legend("Target Prev"))+
  geom_errorbar(aes(ymin=mean, ymax=mean+sd), width=.2,
                position=position_dodge(.9))+ 
  geom_text(
    aes(label = round(mean,2), y = mean+sd+0.015),
    position = position_dodge(0.9),
    vjust = 0, size=3
  )
```
